---
title: '**VISUALIZATION OF PERU COVID-19 TIME SERIES**'
author: "MARCO ARELLANO B. <br> [Twitter: marellanob93](https://twitter.com/marellanob93) <br> [Github: marellanob](https://github.com/marellanob)"
date: "3/31/2021"
output:
  html_document: 
    theme: spacelab
  always_allow_html: yes
  pdf_document: default
subtitle: '**SOURCE: MINSA-PERU / CSSE-JOHN HOPKINS UNIVERSITY**'
link-citations: yes
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      message = FALSE,
                      warning=FALSE)
```

**Keywords: Visualization, R coding, dygraph, plotly, manipulateWidget, Time Series, Covid-19, Peru.**

<style>
/* resize the widget container */
.manipulateWidget { 
  width: 200% !important;
}

/* center the widget */
div.svg-container {
  margin: auto !important;
}
</style>

### Introduction

Hello! I've been learning R programming with RStudio and this is my first visualization project to complete, where I show what I've been learning regarding data visualization. I have studied communication, with experience in film production, and have been drawn to the field of analytics through the area of visualization. I am amazed by how much information can be contained in a single graph. Visualization is like a good movie, show the story and at the same time generates conversations and analysis. 


This project is about the evolution of Covid-19 in Peru from March 6, 2020 to March 31, 2021. Peru, located in South America, has a population of 33 million, and three distinct geographical regions: desertic west coast, central mountain high Andean region, and eastern tropical Amazon range. For more Peru information            <a href="https://www.britannica.com/place/Peru">click here</a>.


Data, presented at national level, correspond to: Cumulative number of cases across daily reports for; a) confirmed cases, b) recovered cases, and c) number of deaths. Since March 15, 2020 Peru has under a series of lockdowns that were relaxed as the number of cases slowed down; or tightened, as the number of cases started to increase in response to the advancement of Covid-19 due to a combination of environmental conditions, health care measures and human behavior response to continuous restrictions that affected mainly, social and economic interactions. An important driver of the early increase of people affected by Covid-19 is the need of a large segment of Peru's population to keep working through the pandemia in very low pay jobs, that made impossible for them to follow government's 'stay in place' regulations.


My visualization objectives are to graph each cumulative series in a single plot, graph each series individual daily cases in a single interactive plot each, and get all these four graphs in a single plot that keeps the interactivity of the individual cases plots. Needlessly to say, I have learned a lot completing these tasks, and enjoyed each step of my learning curve.


### Data

Datasets were downloaded from the github repository for <a href="https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series">CSSE-John Hopkins University</a>. Each dataset contains information at national level, 274 rows for countries in the confirmed and deaths data set; and 259 rows in the recovered dataset, with  data for each day  corresponding to a column, for a total of 439 columns. Date in format *mm-dd-yyyy* gives name to each column. Each country correspond to a row, except for Australia, Canada, China, Denmark, France, Netherlands and UK , that present information at province level. Last access was on 03-31-2021.

Downloaded datasets were in wide format and have the string of characters 'X.' in front of the date giving name to the columns. I've used the `base::substring` command to eliminate these strings. Next, in each dataset, I've added, with the command `dplyr::inner_join`, the variable `continents`, which corresponds to the continent each country is located.

Then, I've converted each dataset from a wide format to a long format using the command `tidyr::pivot_longer`.
As a result the datasets have 5 columns, and 67080 rows. Columns correspond to *country.Region*, *continent*, *region*, *dates*, and *confirmed* (for **confirmed** cases dataset), *recovered* (for **recovered** cases dataset), or *deaths* (for number of **deaths** dataset). Finally, the variable *dates* was converted (from integer) to dates with the command `lubridate::mdy`.

### Technical Specifications

This project was made with:

 - R 4.0.3 GUI 1.73 Catalina build (7892). 
 - R studio Version 1.4.1103 for MacOS.  
 - Mac Book Pro- MacOS High Sierra Version 10.13.6


### R Packages

```{r libraries}
library(readxl)
library(tidyverse)
library(magrittr)
library(lubridate)
library(dplyr)
library(xts)
library(tsibble)
library(slider)
library(dygraphs)
library(plotly)
library(manipulateWidget)
library(knitr)
```


Read Cumulative Confirmed Cases

```{r data_confirmed}
confirmed <- read.csv("/Users/marcoarellano/Desktop/DATA SCIENCE/Covid 19/03.31.2021/DATA/time_series_covid19_confirmed_global.csv")
n_colsc <- dim(confirmed)[2]
n_rowsc <- dim(confirmed)[1]
names(confirmed)[5:n_colsc] <- substring(names(confirmed)[5:n_colsc],2)
tail(confirmed[, 1:6])
```

`confirmed dataset has `r n_rowsc` rows and `r n_colsc` columns.`

Read Cumulative Recovered Cases

```{r data_recovered}
recovered <- read.csv("/Users/marcoarellano/Desktop/DATA SCIENCE/COVID 19/03.31.2021/DATA/time_series_covid19_recovered_global.csv")
n_colsr <- dim(recovered)[2]
n_rowsr <- dim(recovered)[1]
names(recovered)[5:n_colsr] <- substring(names(recovered)[5:n_colsr],2)
tail(recovered[, 1:6])
```

`recovered dataset has `r n_rowsr` rows and `r n_colsr` columns.`

Read Cumulative Deaths Cases

```{r data_deaths}
deaths <- read.csv("/Users/marcoarellano/Desktop/DATA SCIENCE/COVID 19/03.31.2021/DATA/time_series_covid19_deaths_global.csv")
n_colsd <- dim(deaths)[2]
n_rowsd <- dim(deaths)[1]
names(deaths)[5:n_colsd] <- substring(names(deaths)[5:n_colsd],2)
tail(deaths[, 1:6])
```

`deaths dataset has `r n_rowsd` rows and `r n_colsd` columns.`

Read Continent and country list

```{r continents}
continents <- read_excel("~/Desktop/DATA SCIENCE/COVID 19/03.31.2021/DATA/continents_Corrected.xlsx")
head(continents)
```


Here I create the long format for each dataset following the steps:

 1. Add *continent* variable to dataset, use `dplyr::inner_join()`
 2. Use `dplyr::pivot_longer` to change having each date as a column, to have each date as a row, within each country.
 3. Define the column *dates* as a **Date variable**.
 4. Group dataset by *Country.Region*, *continent*, *region*, *dates*.
 5. Get cumulated values of *confirmed* (or *recovered*, or *deaths*) for cases where there are more than one row per country.
 6. Cancel the grouping.
 

Create the long format for accumulated confirmed cases (**confirmed_long**)

```{r confirmed_long}
confirmed_long <- confirmed %>%
  inner_join(continents, by = "Country.Region") %>%
  pivot_longer (
    cols = !c(Province.State, Country.Region, Lat, Long, continent, region),
    names_to = c("dates"),
    values_to = "confirmed") %>%
  mutate(dates = mdy(dates)) %>%
  group_by(Country.Region, continent, region, dates) %>%
  summarise(confirmed = sum(confirmed)) %>%
  ungroup()
n_colscl <- dim(confirmed_long)[2]
n_rowscl <- dim(confirmed_long)[1]
tail(confirmed_long)
```

`confirmed_long dataset has `r n_rowscl` rows and `r n_colscl` columns.`

Create the long format for accumulated recovered cases (**recovered_long**)


```{r recovered_long}
recovered_long <- recovered %>%
  inner_join(continents, by = "Country.Region") %>%
  pivot_longer (
    cols = !c(Province.State, Country.Region, Lat, Long, continent, region),
    names_to = c("dates"),
    values_to = "recovered") %>%
  mutate(dates = mdy(dates))%>%
  group_by(Country.Region, continent, region, dates) %>%
  summarise(recovered = sum(recovered)) %>%
  ungroup()
n_colsrl <- dim(recovered_long)[2]
n_rowsrl <- dim(recovered_long)[1]
tail(recovered_long)
```

`recovered_long dataset has `r n_rowsrl` rows and `r n_colsrl` columns.`

Create the long format for accumulated deaths (**deaths_long**)

```{r deaths_long}
deaths_long <- deaths %>%
  inner_join(continents, by = "Country.Region") %>%
  pivot_longer (
    cols = !c(Province.State, Country.Region, Lat, Long, continent, region),
    names_to = c("dates"),
    values_to = "deaths") %>%
  mutate(dates = mdy(dates)) %>%
  group_by(Country.Region, continent, region, dates) %>%
  summarise(deaths= sum(deaths)) %>%
  ungroup()
n_colsdl <- dim(deaths_long)[2]
n_rowsdl <- dim(deaths_long)[1]
tail(deaths_long)
```

`deaths_long dataset has `r n_rowsdl` rows and `r n_colsdl` columns.`

Below, I create a new column that corresponds to the number of daily cases in each dataset.

Daily cases are achieved by subtracting the cases from day (j-1) to day (j), the difference give us the case increase in a single day. To achieve this, I use the `dplyr::lag` command that allows to find the number of cases in the day before. Using the `lag()` option **default=0**, means that the `lag` value for the first observation will be the same as the observed value for that day.

 1. Order sequentially dates from first date, 01-22-2020,  to final date, 03-31-2021.
 2. Group rows by *Country.Region*.
 3. Create variable *confirmed_dailycases*, or *recovered_dailycases*, or *deaths_dailycases*.
 4. Cancel the grouping within each dataset.

 
Create the variable **confirmed_dailycases** 

```{r lag_confirmed}
confirmed_long <- confirmed_long %>%
  arrange(dates) %>%
  group_by(Country.Region) %>%
  mutate(confirmed_dailycases = confirmed - lag(confirmed, default = 0)) %>%
  ungroup()
tail(confirmed_long)
```

Create the variable **recovered_dailycases** 

```{r lag_recovered}
recovered_long <- recovered_long %>%
  arrange(dates) %>%
  group_by(Country.Region) %>%
  mutate(recovered_dailycases = recovered - lag(recovered, default = 0)) %>%
  ungroup()
tail(recovered_long)
```

Create the variable **deaths_dailycases** 

```{r lag_deaths}
deaths_long <- deaths_long %>%
  arrange(dates) %>%
  group_by(Country.Region) %>%
  mutate(deaths_dailycases = deaths - lag(deaths, default = 0)) %>%
  ungroup()
tail(deaths_long)
```


Here I select data from my country, Peru, following the steps,

 1. Use `dplyr::filter`  to select Peru data in each dataset.
 2. Select variables to be included in each Peru dataset: **Peru_confirmed**, **Peru_recovered**, **Peru_deaths**.
 3. Eliminate the column *Country.Region* because the three datasets refers to Peru.
 4. Use `dplyr::full_join` to create a single dataset with all three series: *confirmed*, *recovered* and *deaths*. The variable *dates* is use for joining the three Peru datasets.
 5. Finally, I create a new variable, *deaths_100k*, which correspond to *(deaths/32625948)10^5)*, as the population of Peru for March 2021 is estimated as <a href="https://www.britannica.com/place/Peru">32 625 948 millions</a>. Values were rounded to one decimal.


```{r Peru}
Peru_confirmed <- confirmed_long %>%
  filter(Country.Region %in% "Peru") %>%
  select(Country.Region, dates, confirmed, confirmed_dailycases)
Peru_confirmed <- Peru_confirmed[,-c(1)]

Peru_recovered <- recovered_long %>%
  filter(Country.Region %in% "Peru") %>%
  select(Country.Region, dates, recovered, recovered_dailycases)
Peru_recovered <- Peru_recovered[,-c(1)]

Peru_deaths <- deaths_long %>%
  filter(Country.Region %in% "Peru") %>%
  select(Country.Region, dates, deaths, deaths_dailycases)
Peru_deaths <- Peru_deaths[,-c(1)]
```

Combine the three datasets in the new dataset **Peru_global** 

```{r Peru_global}
Peru_global <- Peru_confirmed %>%
  full_join(Peru_recovered, by = "dates") %>%
  full_join(Peru_deaths, by = "dates") %>%
  mutate(deaths_100k = ceiling((deaths/32625948)*10^5))
n_colspg <- dim(Peru_global)[2]
n_rowspg <- dim(Peru_global)[1]
tail(Peru_global)
```

`Peru_global dataset has `r n_rowspg` rows and `r n_colspg` columns.`

Next, I create a 7-day Rolling Average variable for *confirmed*,  *recovered*, and *deaths* variables in dataset **Peru_global**.

First, transform **Peru_global** dataset to a time series object. I use the command `tsibble::as_tsibble`


```{r Peru_global_transform}
nr <- nrow(Peru_global)
Peru_global$rid <- seq(1, nr ,1)
Peru_global_ts <- as_tsibble(Peru_global, 
                             key = rid,
                             index = dates)
```


The 7-day rolling average takes seven consecutive values and calculate their average, this average is paired with the central date of the 7-day interval, which correspond to the 4th date, the following 7-day interval is created dropping the earliest date of the interval and adding the next date after the latest date of the interval.

- The 7-day rolling average is created with the command `slider::slide_index_dbl`.  
- The position of our 7-day Rolling Average corresponds to day 4 of consecutive 7-day intervals.
- The 7-day Rolling Averages smooth the day to day observed variation in 7 consecutive days.


```{r 7_DAY_ROLLING AVG}
Peru_global_ts <- Peru_global_ts %>% 
  filter_index("2020-03-06" ~ .) %>% 
  mutate(confirmed7_dailycases = slide_index_dbl(.i = dates,
                                                 .x = confirmed_dailycases,
                                                 .f = mean,
                                                 .before = 3,
                                                 .after= 3),
          recovered7_dailycases = slide_index_dbl(.i = dates,
                                                  .x = recovered_dailycases,
                                                  .f = mean,
                                                  .before = 3,
                                                  .after = 3),
         deaths7_dailycases = slide_index_dbl(.i = dates,
                                              .x = deaths_dailycases,
                                              .f = mean,
                                              .before = 3,
                                              .after = 3),
         confirmed7 = slide_index_dbl(.i = dates,
                                      .x = confirmed,
                                      .f = mean,
                                      .before = 3,
                                      .after = 3),
         recovered7 = slide_index_dbl(.i = dates,
                                      .x = recovered,
                                      .f = mean,
                                      .before = 3,
                                      .after = 3),
         deaths7 = slide_index_dbl(.i = dates,
                                   .x = deaths,
                                   .f = mean,
                                   .before = 3,
                                   .after = 3))
head(Peru_global_ts)
print(Peru_global_ts)
```


## Now let's graph!

I use the `dygraph` library to graph an interactive time series of *confirmed*, *recovered* and *deaths* daily cases. Each plot has 2 variables: the daily number of cases and 7-day rolling average.

The interactive graph allows to zoom in selected time intervals for a more detailed view of the series. 

First Graph is for the Number of Daily Confirmed Cases.


<style>
.html-widget {
    margin: auto;
}
</style>


```{r CONFIRMED_DAILY, fig.align = 'center'}
peru_int_confirmeddaily <- cbind(Peru_global_ts[c(1, 3, 10)])
peru_int_confirmeddaily$confirmed7_dailycases <- round(peru_int_confirmeddaily$confirmed7_dailycases, 0)
rownames( peru_int_confirmeddaily) <- as.POSIXlt( peru_int_confirmeddaily[, 1])
ts_peru_int_confirmeddaily <-  peru_int_confirmeddaily[, -1]
dygraph(ts_peru_int_confirmeddaily, 
        main = "Confirmed Covid-19 Daily cases") %>%
  dySeries("confirmed_dailycases", stepPlot = TRUE, 
           fillGraph = TRUE, color = "lightblue", label = "Confirmed Daily Cases") %>%
  dySeries("confirmed7_dailycases", drawPoints = TRUE, 
           pointShape = "square", color = "darkblue", label = "Rolling Avg 7") %>%
  dyRangeSelector(height = 20) %>%
  dyLegend(width = 300)
```

<br/><br/>
Second Graph is for the Number of Daily Recovered Cases.

```{r DYGRAPGH_RECOVERY_DAILY, fig.align='right'}
peru_int_recovereddaily <- cbind(Peru_global_ts[c(1, 5, 11)])
peru_int_recovereddaily$recovered7_dailycases <- round(peru_int_recovereddaily$recovered7_dailycases, 0)
rownames( peru_int_recovereddaily) <- as.POSIXlt( peru_int_recovereddaily[, 1])
ts_peru_int_recovereddaily <-  peru_int_recovereddaily[, -1]
dygraph(ts_peru_int_recovereddaily, 
        main = " Recovered Covid-19 Daily cases") %>%
  dySeries("recovered_dailycases", stepPlot = TRUE, 
           fillGraph = TRUE, color = "turquoise", label = "Recovered Daily Cases") %>%
  dySeries("recovered7_dailycases", drawPoints = TRUE, 
           pointShape = "circle", color = "green", label = "Rolling Avg 7") %>%
  dyRangeSelector(height = 20) %>%
  dyLegend(width = 300) 
```


<br/><br/>
Third Graph is for the Number of Daily Deaths.


```{r Deaths_Daily}
peru_int_deathsdaily <- cbind(Peru_global_ts[c(1, 7, 12)])
peru_int_deathsdaily$deaths7_dailycases <- round(peru_int_deathsdaily$deaths7_dailycases, 0)
rownames( peru_int_deathsdaily) <- as.POSIXlt( peru_int_deathsdaily[, 1])
ts_peru_int_deathsdaily <-  peru_int_deathsdaily[, -1]
dygraph(ts_peru_int_deathsdaily,
        main = "Covid-19 Daily Deaths") %>%
  dySeries("deaths_dailycases", stepPlot = TRUE, fillGraph = TRUE, 
           color = "orange", label = "Deaths Daily Cases") %>%
  dySeries("deaths7_dailycases", drawPoints = TRUE, pointShape = "square",
           color = "red", label = "Rolling Avg 7") %>%
  dyRangeSelector(height = 20) %>%
  dyLegend(width = 285)
```


<br/><br/>
The `plotly` library is used to create the last graph. For this graph we use the accumulated values and the 7-day rolling average of our 3 variables: *confirmed*, *recovered* and *deaths*.

This chart has 2 y-axis. The y-axis on the left corresponds to the values of confirmed and recovered cases; on the other hand, the right y-axis corresponds to the values of deaths. I considered to have two y-axis because the confirmed and recovered values have a similar range, in contrast to the death values that had a lower range. For that reason, in order to visualize the trend in a better way, it was decided to add the second y-axis.

<style>
.html-widget {
    margin: auto;
}
</style>


```{r, fig.asp=1, fig.height=7, fig.align='center'}
plot_ly() %>%
  add_trace(x = ~Peru_global_ts$dates, y = ~ round(Peru_global_ts$confirmed7, 0), name = "Confirmed",  
            type = 'scatter', mode = 'lines', line = list(color = 'blue', size = 4),
            hoverinfo = "text",
            text = ~paste("Date: ", Peru_global_ts$dates,
                          "<br>",
                          "Confirmed: ", round(Peru_global_ts$confirmed7, 0))) %>%
  add_trace(x = ~Peru_global_ts$dates, y = ~round(Peru_global_ts$recovered7, 0), name = "Recovered",
            type = 'scatter', mode = 'lines', line = list(color = 'green', size = 4),
            hoverinfo = "text",
            text = ~paste("Date: ", Peru_global_ts$dates,
                          "<br>",
                          "Recovered: ", round(Peru_global_ts$recovered7, 0))) %>% 
  add_trace(x = ~Peru_global_ts$dates, y = ~round(Peru_global_ts$deaths7, 0), name = "Deaths", yaxis = "y2",
            type = 'scatter', mode = 'lines', line = list(color = 'red', size = 4),
            hoverinfo = "text",
            text = ~paste("Date: ",Peru_global_ts$dates,
                          "<br>",
                          "Deaths: ", round(Peru_global_ts$deaths7, 0))) %>% 
  layout( title = list(text ="7-day Rolling Average Cumulative Covid-19 cases Peru 2020-2021",
                       size = 10),
          yaxis2 = list(tickfont = list(color = "red"),
                        overlaying = "y",
                        side = "right",
                        title = "Cumulative Deaths",
                        showgrid = FALSE),
          xaxis = list(title = "Dates",
                       color = "black"),
          yaxis = list(tickangle = 0,
                      title = "Cumulative Confirmed and Recovered <br><br><br>",
                      standoff = 90,
                      showgrid = FALSE),
          legend = list(orientation = "h",   
                        xanchor = "center",  
                        x = 0.5,             
                        y = -0.2),
          autosize = T,
          margin = list(l = 100, r = 100, b = 100, t = 100, pad = 20))
```


After all our graphics are ready, I combine them in a single figure that has two columns; the left column has a combined cumulative cases series graph, and the right column has three individual graphs corresponding to daily cases.

I use the command `manipulateWidget::combineWidget`. This command allows to join our interactive graphics in a single image in a quick and easy way.

First, I create a function to combine in a single graph the three cumulative series.


```{r CULUMATES_PLOT_functions}
cumulates_plotly <- function(id){plot_ly() %>%
  add_trace(x = ~Peru_global_ts$dates, y = ~ round(Peru_global_ts$confirmed7, 0), name = "Confirmed",  
            type = 'scatter', mode = 'lines', line = list(color = 'blue', size = 4),
            hoverinfo = "text",
            text = ~paste("Date: ", Peru_global_ts$dates,
                          "<br>",
                          "Confirmed: ", round(Peru_global_ts$confirmed7, 0))) %>%
  add_trace(x = ~Peru_global_ts$dates, y = ~round(Peru_global_ts$recovered7, 0), name = "Recovered",
            type = 'scatter', mode = 'lines', line = list(color = 'green', size = 4),
            hoverinfo = "text",
            text = ~paste("Date: ", Peru_global_ts$dates,
                          "<br>",
                          "Recovered: ", round(Peru_global_ts$recovered7, 0))) %>% 
  add_trace(x = ~Peru_global_ts$dates, y = ~round(Peru_global_ts$deaths7, 0), name = "Deaths", yaxis = "y2",
            type = 'scatter', mode = 'lines', line = list(color = 'red', size = 4),
            hoverinfo = "text",
            text = ~paste("Date: ", Peru_global_ts$dates,
                          "<br>",
                          "Deaths: ", round(Peru_global_ts$deaths7, 0))) %>% 
  layout( title = list(text = "7-day Rolling Average Cumulative Covid-19 cases Peru 2020-2021",
                       size = 10),
          yaxis2 = list(tickfont = list(color = "red"),
                        overlaying = "y",
                        side = "right",
                        title = "Cumulative Deaths",
                        showgrid= FALSE),
          xaxis = list(title = "Dates",
                       color = "black"),
          yaxis = list(tickangle =0,
                      title = "Cumulative Confirmed and Recovered <br><br><br>",
                      standoff = 90,
                      showgrid = FALSE),
          legend = list(orientation = "h",   
                        xanchor = "center",  
                        x = 0.5,             
                        y = -0.2),
          autosize = T,
          margin = list(l = 100, r = 100, b = 100, t = 100, pad = 20))}
```


Second, I create a function to define each component of the right column in the final figure.


```{r DYGRAPH_functions}
c1 <-function(id){dygraph(ts_peru_int_confirmeddaily, 
        main = "Confirmed Covid-19 Daily cases") %>%
  dySeries("confirmed_dailycases", stepPlot = TRUE, 
           fillGraph = TRUE, color = "lightblue", label = "Confirmed Daily Cases") %>%
  dySeries("confirmed7_dailycases", drawPoints = TRUE, 
           pointShape = "square", color = "darkblue", label = "Rolling Avg 7") %>%
  dyRangeSelector(height = 20) %>%
  dyLegend(width = 300)}

r1<-function(id){dygraph(ts_peru_int_recovereddaily, 
        main = " Recovered Covid-19 Daily cases") %>%
  dySeries("recovered_dailycases", stepPlot = TRUE, 
           fillGraph = TRUE, color = "turquoise", label = "Recovered Daily Cases") %>%
  dySeries("recovered7_dailycases", drawPoints = TRUE, 
           pointShape = "circle", color = "green", label = "Rolling Avg 7") %>%
  dyRangeSelector(height = 20) %>%
  dyLegend(width = 300)}

d1<-function(id){dygraph(ts_peru_int_deathsdaily,
        main = "Covid-19 Daily Deaths") %>%
  dySeries("deaths_dailycases", stepPlot = TRUE, fillGraph = TRUE, 
           color = "orange", label = "Deaths Daily Cases") %>%
  dySeries("deaths7_dailycases", drawPoints = TRUE, pointShape = "square",
           color = "red", label = "Rolling Avg 7") %>%
  dyRangeSelector(height = 20) %>%
  dyLegend(width = 285)}
```


To conclude, I use `combineWidget` to arrange the charts in the final figure Additionally, I create the function `write_alt_text` to add an alternative text to the graph.


```{r COMBINE_PLOT,fig.height=8, fig.width=12, fig.align="left"}
write_alt_text <- function(
  chart_type, 
  type_of_data, 
  reason, 
  source){glue::glue(
    "{chart_type} of {type_of_data} where {reason}.<br> \n\nData source from {source}")}

combineWidgets(
  ncol = 2, colsize = c(2,1),
  cumulates_plotly(1),
  title = "Covid-19 Peru Interactive Time Series",
  footer = write_alt_text(
  "<br/>Time Series", 
  "confirmed, recovered and deaths cases from Covid-19 in Peru", 
  "information about the evolution of Covid-19 is needed", 
  "MINSA-Peru/ CSSE-John Hopkins University.<br>Made by Marco Arellano B. Twitter: marellanob93, Github: marellanob") ,
  combineWidgets(
    ncol = 1,
    c1(2),
    r1(3),
    d1(4)))
```
<br/><br/><br/>
<style>
.html-widget {
    margin: left;
}
</style>
